### 2023-01-17 src/reopt.Rmd
read in reopt data

---
title: "reOpt Data"
author: "Grace Bianchi"
date: "2023-01-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(janitor)
library(googlesheets4)
```

```{r}
# read in data
reopt_data <- read_sheet("https://docs.google.com/spreadsheets/d/1VROhOTWY4jSeI_zNgFui4ie6QiIz7uzzli8rF72fAyg/edit#gid=0") %>% 
  clean_names()

```

```{r}
Annual Renewable Electricity (%)
Lifecycle Total tons CO2
Internal Rate of Return (IRR) (%)
Lifecycle Costs of Health Emissions ($)

```