### 2023-01-18 src/climate_risk.Rmd
new NRI climate data and climate risk rmd

---
title: "Climate Risk"
author: "Julia Bickford"
date: '2023-01-18'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
library(janitor)
```
```{r}
#read in National Risk Index (NRI) county data

climate_risk <- read.csv(here("data", "raw", "NRI_Table_Counties.csv")) %>% 
  clean_names()

#filter data by county

```