### 2023-01-19 src/climate_risk.Rmd

---
title: "Climate Risk"
author: "Julia Bickford"
date: '2023-01-18'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(janitor)
library(googlesheets4)
```

```{r}
# read in reopt data (for STCOFIPS codes)
reopt_data <- read_sheet("https://docs.google.com/spreadsheets/d/1VROhOTWY4jSeI_zNgFui4ie6QiIz7uzzli8rF72fAyg/edit#gid=0") %>% 
  clean_names() %>% 
  drop_na()

#read in National Risk Index (NRI) county data

climate_risk <- read_sheet("https://docs.google.com/spreadsheets/d/1J8-PrtmSASmcbLp7Z2TDYsqx1_sSM6naNYPzyMeElGM/edit?usp=sharing")

```