### 2023-01-20 src/climate_risk.Rmd
NRI scores calculated with stcofips id and cbsa number

---
title: "Climate Risk"
author: "Julia Bickford"
date: '2023-01-18'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(janitor)
library(googlesheets4)
```

```{r}
# read in reopt data (for STCOFIPS codes)
reopt_data <- read_sheet("https://docs.google.com/spreadsheets/d/1VROhOTWY4jSeI_zNgFui4ie6QiIz7uzzli8rF72fAyg/edit#gid=0") %>% 
  clean_names() %>% 
  drop_na()

#read in National Risk Index (NRI) county data

climate_risk <- read_sheet("https://docs.google.com/spreadsheets/d/1J8-PrtmSASmcbLp7Z2TDYsqx1_sSM6naNYPzyMeElGM/edit?usp=sharing")

```

```{r}
#pull stcofips numbers for anchor cities
anchor_stcofips <- reopt_data %>% pull(stcofips_5_digit)
```

```{r}
#filter NRI data to only include anchor cities
anchor_climate_risk <- climate_risk %>% filter(STCOFIPS %in% anchor_stcofips)
```

```{r}
#select stcopfips number and overall risk score from NRI data 
anchor_climate_risk_score <- anchor_climate_risk %>% select(STCOFIPS, RISK_SCORE)
```

```{r}
#select stcofips number and cbsa number from from reopt anchor city data
reopt_data_cbsa_stcofips <- reopt_data %>% select(cbsa, stcofips_5_digit)

#rename stcofips col to merge data tables by stcofips number
reopt_data_cbsa_stcofips <- reopt_data_cbsa_stcofips %>% rename("STCOFIPS" =  "stcofips_5_digit")
```

```{r}
#merge data to create dable with: CBSA, stcofips, and risk score
climate_risk_cbsa_stcofips <- merge(reopt_data_cbsa_stcofips, anchor_climate_risk_score, by="STCOFIPS")
```

```{r}
#normalize climate risk scores
climate_risk_cbsa_stcofips <- climate_risk_cbsa_stcofips %>% 
  # normalize climate risk
  mutate(normalized_climate_risk = (RISK_SCORE - min(RISK_SCORE)) / (max(RISK_SCORE) - min(RISK_SCORE)))

#invert normalized scores so that places with less climate risk have a better score
climate_risk_cbsa_stcofips <- climate_risk_cbsa_stcofips %>% 
  # normalize climate risk
  mutate(normalized_inverted_climate_risk = 1-normalized_climate_risk)

```

```{r}
#write result to google sheet
sheet_write(climate_risk_cbsa_stcofips, ss = "https://docs.google.com/spreadsheets/d/1HZCeNGf61RJ9GJs3cLJcl9aNpvmPMBF1Yoss2fytltc/edit?usp=sharing")
```