### 2023-01-23 src/landlord.Rmd

---
title: "landlord"
author: "Virginia Pan"
date: "10/21/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(googlesheets4)
library(janitor)
library(tidyverse)
library(kableExtra)
library(DT)
library(ggplot2)

MSA_landlord_data <- read_sheet("https://docs.google.com/spreadsheets/d/1zCePYRyJ5lb1Z5jvCujg17EtwqpxJN1E1_HwJL66uLM/edit#gid=122953884")

# select_MSAs <- read_sheet("https://docs.google.com/spreadsheets/d/1lmTpSDwVANxdAg5sW87Q8l_gtig7M0EjPPJ6J5_7dwg/edit?usp=sharing")
```

```{r}
#filter by new MSA list
# new_MSA_list <- unique(select_MSAs$cbsa)
# 
# MSA_landlord_data <- MSA_landlord_data %>% 
#   filter(cbsa %in% c(new_MSA_list))
```

```{r}
#security_deposit_limit
MSA_landlord_data <- MSA_landlord_data %>% 
  mutate(
    security_depopsit_limit_num = case_when(
      security_deposit_limit == "N" ~ 1.0, 
      security_deposit_limit == "1_month_rent" ~ 0.0,
      security_deposit_limit == "2_months_rent" ~ 0.5))

# #notice_of_entry_days
# MSA_landlord_data <- MSA_landlord_data %>% 
#   mutate(
#     notice_for_entry_days_num = case_when(
#       notice_for_entry_days == "N" ~ 1.0, 
#       notice_for_entry_days == "0" ~ 1.0,
#       notice_for_entry_days == "0.5" ~ 0.75,
#       notice_for_entry_days == "1" ~ 0.5,
#       notice_for_entry_days == "2" ~ 0.0,))
```

```{r}
#normalize eviction notice (days)
MSA_landlord_data <- MSA_landlord_data %>% 
  mutate(
    eviction_notice_days_normalized = (eviction_notice_days-min(eviction_notice_days))/(max(eviction_notice_days)-min(eviction_notice_days)) 
  )

#"invert" normalized score for eviction notice days
#1 - normalized score (since more eviction notice is worse for landlords)
MSA_landlord_data <- MSA_landlord_data %>% 
  mutate(
    eviction_notice_days_normalized_inverted = 1 - eviction_notice_days_normalized
  )
```

```{r}
#calculate landlord category score

#input landlord weights for each indicator in the following order: eviction notice days, security deposit limit, notice for entry

#Weights must sum to 1
landlord_weights <- c(0.8, 0.2)

#make landlord_score column
MSA_landlord_data <- MSA_landlord_data %>% 
  mutate(
    landlord_score = MSA #this is just a place holder
  )
size <- dim(MSA_landlord_data)
for (i in 1:size[1]) {
  MSA_x_data <- c(MSA_landlord_data$eviction_notice_days_normalized_inverted[i],
                  MSA_landlord_data$security_depopsit_limit_num[i]
                  )
  MSA_landlord_data$landlord_score[i] = weighted.mean(MSA_x_data, landlord_weights)
}
```

```{r}
#If there is rent control, immediately disqualify location (set score to -999)
for (j in 1:size[1]) {
  if(MSA_landlord_data$rent_control[j] == "Y"){
    MSA_landlord_data$landlord_score[j] = -999
  }
}

datatable(MSA_landlord_data, extensions = 'Buttons', options = list(
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel', 'pdf', 'print')
  )
)
```

```{r}
#write result to google sheet
#gs4_create("MSA_landlord_score")
sheet_write(MSA_landlord_data, ss = "https://docs.google.com/spreadsheets/d/1nLpDhM88a3rEBDVK5RZlk2B6QuwVIwvsfE5HeBz0mLo/edit?usp=sharing", sheet = "MSA_landlord_data")
```

```{r}
# mean ("mean_metric"), 95th percentile ("perc_metric"), and distribution for sensitivity analysis

#Eviction Notice (normalized, inverted)
mean_eviction <- mean(MSA_landlord_data$eviction_notice_days_normalized_inverted)
mean_eviction

perc_eviction <- quantile(MSA_landlord_data$eviction_notice_days_normalized_inverted, probs = 0.95)
perc_eviction

eviction_distribution <- density(MSA_landlord_data$eviction_notice_days_normalized_inverted)
plot(eviction_distribution)

#Security Deposit Limit
mean_deposit <- mean(MSA_landlord_data$security_depopsit_limit_num)
mean_deposit

perc_deposit <- quantile(MSA_landlord_data$security_depopsit_limit_num, probs = 0.95)
perc_deposit

deposit_distribution <- density(MSA_landlord_data$security_depopsit_limit_num)
plot(deposit_distribution)

```