### 2023-01-30 src/data_viz.Rmd
started bar chart

---
title: "Data Visualization"
author: "Grace Bianchi"
date: "2023-01-27"
output: html_document
---

```{r setup, include=FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(googlesheets4)
library(tidyverse)
library(janitor)
library(ggplot2)

```

```{r}
total_score_data <- read_sheet("https://docs.google.com/spreadsheets/d/1yqjhJvXUcEiC3qiYWkKNlF6NNpYXWg15zZHwmwpS5Q0/edit#gid=1307699202") %>% 
  rename(financial_score = normalized_irr,
         climate_impact = inverted_n_total_co2,
         health_impact = inverted_n_health_emissions_cost,
         electricity = normalized_annual_renewable,
         climate_risk = normalized_inverted_climate_risk)

real_estate_data <- read_sheet("https://docs.google.com/spreadsheets/d/1Ssmb7pUd4NFZocVp9C75SRMOV6tAJCCt0TjrCYBB8pY/edit#gid=1038418521")

rental_policy_data <- read_sheet("https://docs.google.com/spreadsheets/d/1nLpDhM88a3rEBDVK5RZlk2B6QuwVIwvsfE5HeBz0mLo/edit#gid=983090810")

climate_risk_data <- read_sheet("https://docs.google.com/spreadsheets/d/1HZCeNGf61RJ9GJs3cLJcl9aNpvmPMBF1Yoss2fytltc/edit")
```

```{r}
score_tidy <- total_score_data %>% 
  pivot_longer(cols = c(real_estate_score, landlord_score, financial_score:climate_risk),
               names_to = "criteria",
               values_to = "criteria_score") %>% 
  filter(city_msa != "Portland") %>% 
  group_by(city_msa)

```

```{r}
ggplot(data = score_tidy, aes(y = city_msa, x = total_score_owen, fill = criteria)) + 
    geom_bar()

ggplot(data = score_tidy, aes(y = city_msa, x = total_score_owen, fill = criteria)) + 
    geom_col(position="stacked")
```