### 2023-04-19 src/MSAtotal_scores.Rmd
updated weighting function to calculate sum of the 7 criteria

---
title: "Total Score"
author: "Grace Bianchi"
date: '2022-11-14'
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
library(janitor)
library(googlesheets4)
library(dplyr)
```

```{r}
# read in the data
real_estate_score_data <- read_sheet("https://docs.google.com/spreadsheets/d/1Ssmb7pUd4NFZocVp9C75SRMOV6tAJCCt0TjrCYBB8pY/edit#gid=1038418521") %>% 
  select(cbsa, real_estate_score)

rental_policy_score_data <- read_sheet("https://docs.google.com/spreadsheets/d/1nLpDhM88a3rEBDVK5RZlk2B6QuwVIwvsfE5HeBz0mLo/edit#gid=983090810") %>% 
  select(cbsa, landlord_score)

reopt_data <- read_sheet("https://docs.google.com/spreadsheets/d/1tPqpmS8JOeRyunEmSyFYPYwshKDGx54Z-MISgYO3wZE/edit#gid=1874679054") %>% 
  select(cbsa, city_msa, state, normalized_irr, inverted_n_total_co2, inverted_n_health_emissions_cost, normalized_annual_renewable)

climate_risk_data <- read_sheet("https://docs.google.com/spreadsheets/d/1HZCeNGf61RJ9GJs3cLJcl9aNpvmPMBF1Yoss2fytltc/edit#gid=2045475301", sheet = "climate_risk+EAL") %>% 
  select(cbsa, normalized_inverted_climate_risk)

```

```{r read in data}
criteria_all <- real_estate_score_data %>% 
  inner_join(rental_policy_score_data)  %>% 
  inner_join(reopt_data) %>% 
  inner_join(climate_risk_data) # combine all criteria scores into one dataframe

### Cleaning up the data order and column names
# reorder columns
col_order <-c("cbsa", "city_msa", "state", "real_estate_score", "landlord_score", "normalized_annual_renewable", "normalized_irr", "inverted_n_total_co2",  "inverted_n_health_emissions_cost", "normalized_inverted_climate_risk")

# rename columns
criteria_all <- criteria_all[, col_order] %>% 
  rename(financial_score = normalized_irr,
         co2_score = inverted_n_total_co2,
         health_impact_score = inverted_n_health_emissions_cost,
         electricity_score = normalized_annual_renewable,
         climate_risk_score = normalized_inverted_climate_risk) %>% 
  mutate(real_estate_score = as.numeric(real_estate_score),
         landlord_score = as.numeric(landlord_score))

```

```{r FINAL student weights}
# weight criteria for students
criteria_weights_students <- c(0.18, 0.0, 0.10, 0.19, 0.17, 0.14, 0.22)

#make landlord_score column
criteria_all <- criteria_all %>% 
  mutate(equity_scenario_score = cbsa,
         client_FINAL_score = cbsa) #this is just a place holder

for (i in 1:nrow(criteria_all)) {
  cbsa_data <- c(criteria_all$real_estate_score[i],
                criteria_all$landlord_score[i],
                 criteria_all$financial_score[i],
                 criteria_all$co2_score[i],
                criteria_all$health_impact_score[i],
                criteria_all$electricity_score[i],
                criteria_all$climate_risk_score[i]
                  )
  criteria_all$equity_scenario_score[i] = weighted.mean(cbsa_data, criteria_weights_students)
}

```

```{r Client FINAL WEIGHTS}

client_weights <- c(0.18, 0.07, 0.11, 0.07, 0.0, 0.07, 0.50)

# multiply column by respective weight and sum for total

for (i in 1:nrow(criteria_all)) {
  cbsa_data <- c(criteria_all$real_estate_score[i],
                criteria_all$landlord_score[i],
                 criteria_all$financial_score[i],
                 criteria_all$co2_score[i],
                criteria_all$health_impact_score[i],
                criteria_all$electricity_score[i],
                criteria_all$climate_risk_score[i]
                  )
  criteria_all$client_FINAL_score[i] = weighted.mean(cbsa_data, client_weights) 
}

```

```{r}

criteria_all <- criteria_all %>% 
  mutate_if(is.numeric,
            round,
            digits = 5)

# creating a function with criteria data and weight inputs
calculate_wt_criteria <- function(data, weights){
    wt_criteria <- data %>%
      # multiply each criteria by respective weight
      mutate(wt_real_estate = real_estate_score * weights[1],
             wt_landlord_policy = landlord_score * weights[2], 
             wt_solar_irr = financial_score * weights[3],
            wt_co2 = co2_score * weights[4], 
             wt_health_impacts = health_impact_score * weights[5],
             wt_electricity = electricity_score * weights[6],
             wt_climate_risk = climate_risk_score * weights[7]) %>% 
      # select weighted columns and rename them
      select(!c(ends_with("_score"))) %>%
      # create column for investment score
      mutate(investment_score = 0) %>%
      # for loop to sum weighted columns for each row
      mutate(investment_score = apply(.[, 4:10], 1, sum)) %>% 
      rename("Real Estate"="wt_real_estate",
             "Landlord Policy"="wt_landlord_policy",
             "Electricity Production"="wt_electricity", 
             "CO2 Emission Reduction"="wt_co2",
             "Climate Risk Avoided"="wt_climate_risk",
             "Health Impacts"="wt_health_impacts", 
             "Solar IRR"="wt_solar_irr") 
 }

#sheet_write(criteria_all, ss = "https://docs.google.com/spreadsheets/d/1yqjhJvXUcEiC3qiYWkKNlF6NNpYXWg15zZHwmwpS5Q0/edit#gid=0", sheet = "FINAL_total_score")
```

### Data Visualization- barcharts

```{r tidy data}
# client weights
wt_criteria_client <- calculate_wt_criteria(criteria_all, client_weights)  %>%  
  unite("msa_state", city_msa:state, sep = ", ")

client_score_tidy <- wt_criteria_client  %>% 
  pivot_longer(cols = c(3:9), # selecting criteria cols
               names_to = "criteria",
               values_to = "wt_criteria_score") %>% 
 filter(msa_state != "Portland, ME")  # removing portland (negative number)

# equity weights
wt_criteria_equity <- calculate_wt_criteria(criteria_all, criteria_weights_students) %>% 
  unite("msa_state", city_msa:state, sep = ", ")

# tidy data for bar chart
equity_tidy <- wt_criteria_equity %>% 
  pivot_longer(cols = c(3:9), # selecting criteria cols
               names_to = "criteria",
               values_to = "wt_criteria_score") 
```

```{r client data viz}
tiff(filename = "client_criteria_barchart.tiff",width = 8, height = 6, units = "in", res = 300, pointsize = 12)
ggplot(data = client_score_tidy, aes(y = reorder(msa_state,  wt_criteria_score, FUN=sum), x = wt_criteria_score)) +
    geom_col(aes(fill = factor(criteria, levels = c("Climate Risk Avoided", "CO2 Emission Reduction", "Electricity Production", "Solar IRR", "Real Estate", "Landlord Policy", "Health Impacts")))) +
  theme_bw() +
  labs(x = "Total Score",
       y = " ",
       title = "") +
  guides(fill=guide_legend(title="Criteria (weighted)")) +
   scale_fill_manual(values = c("palevioletred1","orange",  "gold", "#44AA99", "dodgerblue4", "#BC80BD", "#6495ED"),
      breaks = c("Climate Risk Avoided", "CO2 Emission Reduction", "Electricity Production", "Solar IRR", "Real Estate", "Landlord Policy", "Health Impacts")) +
        scale_x_continuous(limits = c(0, 0.8), expand = c(0,0), breaks=seq(0, 0.8, 0.1)) 
#+ theme(legend.position = "none")
dev.off()
```

```{r student data viz}
# ## read in weighted criteria sheet
# student_weighted_criteria <- read_sheet(ss = "https://docs.google.com/spreadsheets/d/1yqjhJvXUcEiC3qiYWkKNlF6NNpYXWg15zZHwmwpS5Q0/edit#gid=0", sheet = "wtd_criteria_stud")

#tiff(filename = "wt_criteria_student_chart.tiff", width = 8, height = 6, units = "in", res = 300)
 ggplot(data = equity_tidy, aes(y = reorder(msa_state, wt_criteria_score, FUN=sum), x = wt_criteria_score)) + 
    geom_col(aes(fill = criteria)) +
  labs(x = "Total Score",
       y = " ",
       title = "") +
  theme_bw() +
  guides(fill=guide_legend(title="Criteria (weighted)")) +
   scale_fill_manual(values = c("palevioletred1","orange",  "gold", "#44AA99", "dodgerblue4", "#BC80BD", "#6495ED"),
      breaks = c("Climate Risk Avoidance", "CO2 Abatement", "Electricity Production", "Solar IRR", "Real Estate", "Landlord Policy", "Health Impacts")) +
  scale_x_continuous(limits = c(0, 1.0), expand = c(0,0), breaks=seq(0, 1.0, 0.1))

#dev.off()
```

```{r}
library(patchwork)

# combined visual of client and equity score results
p1 <- ggplot(data = client_score_tidy, aes(y = reorder(msa_state,  wt_criteria_score, FUN=sum), x = wt_criteria_score)) +
    geom_col(aes(fill = criteria)) +
  labs(x = "Total Score",
       y = " ",
       title = "") +
  theme_bw() +
  guides(fill=guide_legend(title="Criteria (weighted)")) +
   scale_fill_manual(values = c("palevioletred1","orange",  "gold", "#44AA99", "dodgerblue4", "#BC80BD", "#6495ED"),
      breaks = c("Climate Risk Avoidance", "CO2 Abatement", "Electricity Production", "Solar IRR", "Real Estate", "Landlord Policy", "Health Impacts")) +
        scale_x_continuous(limits = c(0, 0.8), expand = c(0,0), breaks=seq(0, 0.8, 0.1)) +
  theme(legend.position = "none") +
  theme(axis.title.y = element_text(size = 20))

p2 <- ggplot(data = equity_tidy, aes(y = reorder(msa_state, wt_criteria_score, FUN=sum), x = wt_criteria_score)) + 
    geom_col(aes(fill = criteria)) +
  labs(x = "Total Score",
       y = " ",
       title = "") +
  theme_bw() +
  guides(shape = guide_legend(override.aes = list(size = 0.5))) +
  guides(fill=guide_legend(title="Criteria (weighted)")) +
   scale_fill_manual(values = c("palevioletred1","orange",  "gold", "#44AA99", "dodgerblue4", "#BC80BD", "#6495ED"),
      breaks = c("Climate Risk Avoidance", "CO2 Abatement", "Electricity Production", "Solar IRR", "Real Estate", "Landlord Policy", "Health Impacts")) +
        scale_x_continuous(limits = c(0, 0.8), expand = c(0,0), breaks=seq(0, 0.8, 0.1)) +
    theme(axis.title.y = element_text(size = 11))

# tiff(filename = "data_vz.tiff", width=12, units = "in", res=100)
 p1/p2 + 
  plot_annotation(tag_levels = 'A') & theme(plot.tag = element_text(size = 16)) 
 ggsave("test_dim_2plots.jpeg", plot = plots, width = 6, height = 4, units = "in", dpi = 300)

 #rstudioapi::savePlotAsImage(file = "myplot.png",width = dev.size(units="px")[[1]],height=dev.size(units="px")[[2]])

```

Figure 1. Ranked metropolitan areas (n=29) based on investment favorability for rooftop solar PV on mutlifamily houses. Areas were scored using six criteria, each given a respective weight given stakeholder priorities. Each criteria is comprised of indicator metrics using data from the U.S. Census Bureau and NREL's reOpt tool. Graph A represents a real estate investor stakeholder who prioritizes climate risk due to long-term investments. B represents a equity-focused stakeholder who values impacts on the community, emissions, and