### 2023-05-04 src/landlord.Rmd
create excel with landlord policies for each state

---
title: "landlord"
author: "Virginia Pan"
date: "10/21/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(googlesheets4)
library(janitor)
library(tidyverse)
library(kableExtra)
library(DT)
library(ggplot2)
library(readxl)
library(here)

MSA_landlord_data <- read_sheet("https://docs.google.com/spreadsheets/d/1zCePYRyJ5lb1Z5jvCujg17EtwqpxJN1E1_HwJL66uLM/edit#gid=122953884")

# for shiny app
state_deposit_data<- read_xlsx(here("data/raw/landlord_state_data.xlsx")) %>% clean_names()

eviction_notice <- read_xlsx(here("data/raw/landlord_state_data.xlsx"), sheet= "Sheet2") %>% 
  clean_names() %>% 
  rename(eviction_days = sel_min_nonpayment)

```

```{r wrangling state deposit limits}
# Define a regular expression pattern to match target phrases
phrase_pattern <- "(one|two)\\smonth|no\\sstatutory\\slimit"

# Define a function to extract matched words from each sentence
extract_words <- function(sentence, pattern) {
  words <- regmatches(sentence, gregexpr(pattern, sentence, ignore.case = TRUE))[[1]]
  if (length(words) > 0) {
    return(paste(words, collapse = " "))
  } else {
    return(NA)
  }
}

state_deposit_data$deposit_limit <- NA

# Create a new column with matched words
state_deposit_data <- state_deposit_data %>%
  mutate(matched_words = sapply(limit, extract_words, pattern = phrase_pattern))

# put matched words in new column
state_deposit_data$deposit_limit <- sapply(state_deposit_data$matched_words, extract_words, pattern = phrase_pattern)

# rename deposit data to apply weights
landlord_state_data <- state_deposit_data %>%
  mutate(deposit_limit = case_when(
    matched_words == "One month"   ~ "1_month_rent",
    matched_words == "Two month" ~ "2_months_rent",
    matched_words == "No statutory limit" ~ "N",
    TRUE ~ NA_character_
  )) %>% 
  select(state, limit, deposit_limit) %>% 
  inner_join(eviction_notice, by = join_by(state == jurisdictions))

#write result to google sheet
sheet_write(landlord_state_data, ss = "https://docs.google.com/spreadsheets/d/1A6o934oBp9Sdbxrx30GkVrGaX9PyzpS1pZc8Pz6Aoro/edit#gid=0", sheet = "MSA_landlord_data") # manually edited eviction data

# write output 
write.xlsx(landlord_state_data, here("data/intermediate/demographic.xlsx"), sheetName = "Sheet1")
```

```{r}
#security_deposit_limit
MSA_landlord_data <- MSA_landlord_data %>% 
  mutate(
    security_depopsit_limit_num = case_when(
      security_deposit_limit == "N" ~ 1.0, 
      security_deposit_limit == "1_month_rent" ~ 0.0,
      security_deposit_limit == "2_months_rent" ~ 0.5))

```

```{r}
#normalize eviction notice (days)
MSA_landlord_data <- MSA_landlord_data %>% 
  mutate(
    eviction_notice_days_normalized = (eviction_notice_days-min(eviction_notice_days))/(max(eviction_notice_days)-min(eviction_notice_days)) 
  )

#"invert" normalized score for eviction notice days
#1 - normalized score (since more eviction notice is worse for landlords)
MSA_landlord_data <- MSA_landlord_data %>% 
  mutate(
    eviction_notice_days_normalized_inverted = 1 - eviction_notice_days_normalized
  )
```

```{r}
#calculate landlord category score

#input landlord weights for each indicator in the following order: eviction notice days, security deposit limit, notice for entry

#Weights must sum to 1
landlord_weights <- c(0.8, 0.2)

#make landlord_score column
MSA_landlord_data <- MSA_landlord_data %>% 
  mutate(
    landlord_score = MSA #this is just a place holder
  )
size <- dim(MSA_landlord_data)
for (i in 1:size[1]) {
  MSA_x_data <- c(MSA_landlord_data$eviction_notice_days_normalized_inverted[i],
                  MSA_landlord_data$security_depopsit_limit_num[i]
                  )
  MSA_landlord_data$landlord_score[i] = weighted.mean(MSA_x_data, landlord_weights)
}
```

```{r}
#If there is rent control, immediately disqualify location (set score to -999)
for (j in 1:size[1]) {
  if(MSA_landlord_data$rent_control[j] == "Y"){
    MSA_landlord_data$landlord_score[j] = -999
  }
}

```

```{r}
#write result to google sheet
#gs4_create("MSA_landlord_score")
sheet_write(MSA_landlord_data, ss = "https://docs.google.com/spreadsheets/d/1nLpDhM88a3rEBDVK5RZlk2B6QuwVIwvsfE5HeBz0mLo/edit?usp=sharing", sheet = "MSA_landlord_data")
```

```{r}
# mean ("mean_metric"), 95th percentile ("perc_metric"), and distribution for sensitivity analysis

#Eviction Notice (normalized, inverted)
mean_eviction <- mean(MSA_landlord_data$eviction_notice_days_normalized_inverted)
mean_eviction

perc_eviction <- quantile(MSA_landlord_data$eviction_notice_days_normalized_inverted, probs = 0.95)
perc_eviction

eviction_distribution <- density(MSA_landlord_data$eviction_notice_days_normalized_inverted)
plot(eviction_distribution)

#Security Deposit Limit
mean_deposit <- mean(MSA_landlord_data$security_depopsit_limit_num)
mean_deposit

perc_deposit <- quantile(MSA_landlord_data$security_depopsit_limit_num, probs = 0.95)
perc_deposit

deposit_distribution <- density(MSA_landlord_data$security_depopsit_limit_num)
plot(deposit_distribution)

```