### 2023-05-09 src/real_estate.Rmd
outputting metric data to excel files

---
title: "Real Estate"
author: "Grace Bianchi"
date: '2022-10-21'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(janitor) # clean_names()
library(googlesheets4)
library(DT)
library(openxlsx)
library(here)
source(here("R/normalize.R"))

city_occupancy_data <- read_sheet("https://docs.google.com/spreadsheets/d/10n55Gt0rySN96OGTQ2_8PzeTcX66n26sOlZBiuiAeBU/edit#gid=808954990") %>% 
  clean_names() %>% 
  rename(occ_rate2021 = percent_occupancy_2021)

MSA_dempgraphic_data <- read_sheet("https://docs.google.com/spreadsheets/d/1lmTpSDwVANxdAg5sW87Q8l_gtig7M0EjPPJ6J5_7dwg/edit?usp=sharing", sheet = "MSA_demographics") %>% 
  clean_names() %>%

  inner_join(city_occupancy_data, by = join_by(cbsa== cbsa, anchor_city==city))

```

```{r}
# calculate metrics
real_estate_metrics<- MSA_dempgraphic_data %>% 
    # calculate rent change
  mutate(med_rent_change = (med_rent2022 - med_rent2021)/med_rent2021 * 100,
         house_price_2021 = as.numeric(house_price_2021),
         income_to_homeprice = med_income/house_price_2021,
         med_rent_to_income = (med_rent2021*12)/med_income) %>%  # multiply monthly rent x 12 for annual rent!
  select(!c(state.y, occ_rate2021, popestimate2021, med_rent2021, med_rent2022, house_price_2021, med_income)) %>% 
  rename(pop_change = pop_percent_change20_21)

write.xlsx(real_estate_metrics, here("data/intermediate/metric_data.xlsx"), sheetName = "real_estate_metrics")

#sheet_write(compliedMSA_realEstate, ss = "https://docs.google.com/spreadsheets/d/1lmTpSDwVANxdAg5sW87Q8l_gtig7M0EjPPJ6J5_7dwg/edit#gid=426650684", sheet = "real_estate_metrics")

```

# Normalize Metrics

```{r}
n_real_estate_metrics <- real_estate_metrics %>%
  mutate(n_pop_change = normalize(pop_change),
         n_employ_change = normalize(employ_change),
         n_income_homprice = 1- normalize(income_to_homeprice), # subtracted from 1 so higher values are more favorable
         n_rent_income = 1 - normalize(med_rent_to_income),
         n_rent_change = normalize(med_rent_change)

        )

```

```{r}
# normalize columns into new data table
n_MSA_realEstate <- compliedMSA_realEstate %>% 
  # normalize population data
  mutate(n_pop_change = (pop_percent_change20_21 - min(pop_percent_change20_21))/(max(pop_percent_change20_21)-min(pop_percent_change20_21))) %>%
  # normalize employment growth
  mutate(n_employ_change = (avg_emp_YoY - min(avg_emp_YoY)) / (max(avg_emp_YoY) - min(avg_emp_YoY))) %>% 
  # normalize rent to income
  mutate(n_med_rent_to_income = (med_rent_to_income- min(med_rent_to_income)) / (max(med_rent_to_income) - min(med_rent_to_income))) %>% 
  # normalize income to home price
  mutate(n_med_rent_change = (med_rent_change - min(med_rent_change)) / (max(med_rent_change) - min(med_rent_change))) %>% 
  mutate(n_income_to_homeprice = (income_to_homeprice - min(income_to_homeprice)) / (max(income_to_homeprice) - min(income_to_homeprice))) %>% 
  mutate(n_avg_occupancy = (occ_rate2021 - min(occ_rate2021)) / (max(occ_rate2021) - min(occ_rate2021)), na.rm = FALSE)

```

```{r}
# invert normalized rent to income ratio so lower values (more favorable) have higher scores
n_MSA_realEstate <- n_MSA_realEstate %>% 
  mutate(n_med_rent_to_income_inverted = 1 - n_med_rent_to_income) %>% 
  mutate(n_income_to_homeprice_inverted = 1 - n_income_to_homeprice)

```

```{r}
# weights from owen
real_estate_weights <- c(0.35, 0.20, 0.15, 0.10, 0.10, 0.10)

#make real_estate_score column
n_MSA_realEstate <- n_MSA_realEstate %>% 
  select(cbsa, MSA, starts_with("n_"), ) %>% 
  mutate(real_estate_score = MSA) 
```

```{r}
size <- dim(n_MSA_realEstate)

for (i in 1:size[1]) {
  MSA_x_data <- c(n_MSA_realEstate$n_pop_change[i],
                  n_MSA_realEstate$n_employ_change[i],
                  n_MSA_realEstate$n_avg_occupancy[i],
                  n_MSA_realEstate$n_med_rent_change[i],
                  n_MSA_realEstate$n_med_rent_to_income_inverted[i], 
                  n_MSA_realEstate$n_income_to_homeprice_inverted[i]
                  )
  n_MSA_realEstate$real_estate_score[i] = sum(MSA_x_data * real_estate_weights)
}

```

```{r}
wt_real_estate_metrics <-n_MSA_realEstate %>%

   mutate(_score = apply(.[, 4:10], 1, sum))
  mutate(weighted_pop_change = 0.35*n_pop_change,
                  weighted_emp_change = 0.20*n_employ_change,
                  weighted_occ = 0.15 * n_avg_occupancy,
                  weighted_rent_change = 0.10*n_med_rent_change,
                  weighted_rent_income = 0.10 * n_med_rent_to_income_inverted, 
                  weighted_income_homeprice = 0.10 * n_income_to_homeprice_inverted) %>% 
  mutate(real_estate_score = weighted_pop_change + weighted_emp_change + weighted_occ + weighted_rent_change + weighted_rent_income + weighted_income_homeprice)

```

```{r}
#write result to google sheet
#gs4_create("anchor_real_estate_score")
sheet_write(n_MSA_realEstate, ss = "https://docs.google.com/spreadsheets/d/1Ssmb7pUd4NFZocVp9C75SRMOV6tAJCCt0TjrCYBB8pY/edit?usp=sharing", sheet = "anchor_realEstate")
```