### 2023-05-12 src/01_metric_normalization.Rmd

---
title: "landlord"
author: "Virginia Pan"
date: "10/21/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(googlesheets4)
library(janitor)
library(tidyverse)
library(DT)
library(readxl)
library(xlsx)
library(here)
source(here("R/normalize.R"))

```

# Calculate and Normalize Real Estate Metrics

```{r}
## read in real estate and demographic data 
city_occupancy_data <- read_sheet("https://docs.google.com/spreadsheets/d/10n55Gt0rySN96OGTQ2_8PzeTcX66n26sOlZBiuiAeBU/edit#gid=808954990") %>% 
  clean_names() %>% 
  rename(occ_rate2021 = percent_occupancy_2021)

anchor_city <- read.xlsx(here("data/intermediate/demographic_data.xlsx"), sheetName="anchor_city") %>% 
  clean_names() %>% 
   inner_join(city_occupancy_data, by = join_by(cbsa== cbsa, anchor_city==city)) %>% 
  select(cbsa, anchor_city, occ_rate2021)

MSA_demographic_data <- read_csv(here("data/intermediate/demographic_data.csv")) %>% 
  clean_names() %>% 
  left_join(anchor_city)

# calculate real estate metrics
real_estate_metrics<- MSA_demographic_data %>% 
    # calculate rent change
  mutate(cbsa = factor(cbsa),
         rent_change = (median_rent2022 - median_rent2021)/median_rent2021 * 100,
         house_price_2021 = as.numeric(house_price_2021),
         income_to_homeprice = median_income/house_price_2021,
         rent_to_income = (median_rent2021*12)/median_income) %>%  # multiply monthly rent x 12 for annual rent!
  select(!c(na, popestimate2021, median_rent2021, median_rent2022, house_price_2021, median_income)) %>% 
  select(cbsa, anchor_city, msa, state, pop_growth, avg_emp_change, 7:10)

## normalization
n_real_estate_metrics <- real_estate_metrics %>%
  # normalize cols and rename columns to start with n
  mutate(across(5:10, normalize, .names = "n_{.col}")) %>% 
  mutate(n_income_to_homeprice = 1- n_income_to_homeprice, # subtracted from 1 so higher values are more favorable
         n_rent_to_income = 1 - n_rent_to_income) %>% 
  select(cbsa, msa, anchor_city, state, starts_with("n_"))

```

# Normalizing Landlord Metrics

```{r}
MSA_landlord_data <- read_csv(here("data/intermediate/landlord_data.csv")) %>% 
  clean_names()

#security_deposit_limit
landlord_metrics <- MSA_landlord_data %>% 
  mutate(
    security_deposit_limit_num = case_when(
      security_deposit_limit == "N" ~ 1.0, 
      security_deposit_limit == "1_month_rent" ~ 0.0,
      security_deposit_limit == "2_months_rent" ~ 0.5)) %>% 
    #"invert" normalized score for eviction notice days
#1 - normalized score (since more eviction notice is worse for landlords)
  mutate(n_eviction_notice = 1- normalize(eviction_notice_days),
         n_security_deposit = normalize(security_deposit_limit_num))

# assign numeric values to security deposit limit for all states
#state_landlord_policy <- state_landlord_policy %>% 
 # mutate(security_deposit_limit = case_when(
  #    deposit_limit == "N" ~ 1.0, 
   #   deposit_limit == "1_months_rent" ~ 0.0,
    #  deposit_limit == "1_month_rent" ~ 0.0,
    #  deposit_limit == "1.5_months_rent" ~ 0.25,
    #  deposit_limit == "2_months_rent" ~ 0.5,
    #  deposit_limit == "2.5_months_rent" ~ 0.75,
    #  deposit_limit == "3_months_rent" ~ 0.75)) %>% 
   #"invert" normalized score for eviction notice days
#1 - normalized score (since more eviction notice is worse for landlords)
  # mutate(n_eviction_notice = 1- normalize(eviction_days),
  #       n_security_deposit = normalize(security_deposit_limit))
```

# reOpt Metric data wrangling and normalization

```{r}
# read in reopt data (for STCOFIPS codes)
reopt_metrics <- read_xlsx(here("data/intermediate/reopt_data_outputs.xlsx"), sheet="reopt_msa_outputs") %>% 
  clean_names() %>% 
  drop_na() %>% 
  rename(
  electricity_generation = annual_renewable_electricity_percent_of_total,
  total_co2 = lifecycle_total_t_co2,
  health_emissions_cost = lifecycle_costs_of_health_emissions_difference_from_bau) %>% 
  select(1:5, electricity_generation, total_co2, health_emissions_cost, irr)

# normalize crit
n_reopt_metrics <- reopt_metrics %>% 
mutate(across(electricity_generation:irr, normalize, .names = "n_{.col}"),
       n_health_emissions_cost = 1-n_health_emissions_cost,
       n_total_co2 = 1 - n_total_co2)

#write_csv(n_reopt_metrics, here("data/intermediate/reopt_metrics.csv"))

```

# reOpt Metric data wrangling and normalization

```{r}
#read in National Risk Index (NRI) county data
climate_risk_metrics <- read_csv(here("data/raw/NRI_Table_Counties.csv")) %>% 
  clean_names() %>% 
  select(state, stateabbrv, county, countytype, risk_score) %>% 
  inner_join(reopt_metrics, by = join_by(stateabbrv==state, county==county)) %>% 
  filter(countytype == "County") # to remove richmond city

# normalize climate risk score
n_climate_risk <- climate_risk_metrics %>% 
  mutate(n_climate_risk = 1- normalize(risk_score)) %>%  #invert normalized scores so that places with less climate risk have a better score
  select(cbsa, city_msa, state, stateabbrv, risk_score, n_climate_risk)

 #write_csv(n_climate_risk, here("data/intermediate/climate_risk_metrics.csv"))
```