### 2023-05-12 src/reopt_api.Rmd

---
title: "Reopt"
author: "Grace Bianchi"
date: "2023-04-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(httr)
library(RJSONIO)
library(jsonlite)
library(openxlsx)
library(writexl)
library(here)
library(tidyverse)
library(janitor)
```

### Obtain MSA city information
Only used once to generate table and output results into data/intermediate folder to decrease runtime for the APi

```{r, eval= FALSE}
token <- "eyJ0eXAiOiJKV1QiLCJhbGciOiJSUzI1NiIsImp0aSI6ImNmMmU0Yzg1ZWZlMDNiYTkxNTA4YzdlNWJiYjE1NjVkNzIyNjg4MzgwYTE4OWNjNmJkOWRjNjFiZWUyZjIwYzg0ZmJjNmZkOTc0MDg4YzRmIn0.eyJhdWQiOiI2IiwianRpIjoiY2YyZTRjODVlZmUwM2JhOTE1MDhjN2U1YmJiMTU2NWQ3MjI2ODgzODBhMTg5Y2M2YmQ5ZGM2MWJlZTJmMjBjODRmYmM2ZmQ5NzQwODhjNGYiLCJpYXQiOjE2ODI0ODQyOTYsIm5iZiI6MTY4MjQ4NDI5NiwiZXhwIjoxOTk4MTAzNDk2LCJzdWIiOiI0ODUyMyIsInNjb3BlcyI6W119.G7NL2Khpo602p_fzk6EiAZ0-KdXsVW-IO9mUqwHem0YFLaTQZezotmLqXmbcTi7BxRzc8UHNYBw415CCzKo81w"

# cbsa-ziip data
cbsa_url <- "https://www.huduser.gov/hudapi/public/usps?type=8&query=All"

# zip-cbsa : type3
#zip_url <- "https://www.huduser.gov/hudapi/public/usps?type=3&query=All"

# set the authorization header
headers <- add_headers("Authorization" = paste("Bearer", token))

# Send the GET request to retrieve the results
cbsa_response <- GET(url = cbsa_url,add_headers("Authorization" = paste("Bearer", token)))

# Extract the response content as JSON and convert it to a list
response_content <- content(cbsa_response, as="text")
response_cbsa <- fromJSON(response_content)

# make data frame
cbsa_city_data <- data.frame("cbsa"= response_cbsa$data$results$cbsa,
                            "city"= response_cbsa$data$results$city,
                            "state"= response_cbsa$data$results$state)  
# output into excel
#write.xlsx(cbsa_city_data, here("data/intermediate/cbsa_city_data.xlsx"), sheetName = "Sheet1")
```

# read in US city with longitude and latitude data from https://simplemaps.com/data/us-cities
```{r}
uscities <- read_csv(here("data/raw/uscities.csv")) %>% 
  mutate(city = tolower(city)) %>% 
  select(state_id, city, lat, lng, population, density, zips) %>% 
  group_by(state_id, city)

city_cbsa_api_data <- read.xlsx(here("data/intermediate/cbsa_city_data.xlsx")) %>% 
   mutate(city = tolower(city)) %>% 
  group_by(state, city)

```

# create cbsa and city data frame that was extracted using hud API
```{r}
cbsa_city_lat_long <- city_cbsa_api_data %>% 
    filter(cbsa != "99999") %>% 
  inner_join(uscities, by = join_by(state == state_id, city == city)) %>%    # removes cities without a cbsa
  distinct()

one_city_msa_data <- cbsa_city_lat_long %>% 
  group_by(cbsa) %>% 
  summarize(max_pop = max(population)) %>% 
  left_join(cbsa_city_lat_long, by=join_by(cbsa== cbsa, max_pop==population), keep=FALSE) %>% 
  # select one zipcode to run in open ei api to get utility data
  separate(col = zips,
           sep = " ",
           into = "zips",
           extra = "drop")

```

# get population estimates for each Metro area in 2021  using census api
```{r polulation estimates}
api_key <- "ebab921b3002df9b71881ad6c426f34281ce0e11"
# Add key to .Renviron
Sys.setenv(CENSUS_KEY="ebab921b3002df9b71881ad6c426f34281ce0e11")
# Reload .Renviron
readRenviron("~/.Renviron")
# Check to see that the expected key is output in your R console
Sys.getenv("CENSUS_KEY")

host <- "https://api.census.gov/data"
year <- "2021"
pop_est_stem <- "pep/population"

job_end <- "acs/acs1?get=NAME,B01001_001E&for=metropolitan%20statistical%20area/micropolitan%20statistical%20area:*&key="

version_url <- file.path(host, year, job_end)
pop_list <- fromJSON(version_url) %>% # retrieved cbsa numbers, msa names, and some value??
  row_to_names(1) %>% clean_names() %>% 
  as.data.frame() %>% 
  rename("msa"= name,
                             "cbsa" = metropolitan_statistical_area_micropolitan_statistical_area,
                             "pop_esta"= b01001_001e)

MSA_w_lat_lng <- pop_list %>% 
  inner_join(cbsa_city_lat_long, by = join_by(cbsa==cbsa))

```

## OpenEI 
```{r}
## open EI api
openEI_api_key <- "rU6MEqhK9OJ7obAnmcbzYhpFOenfBM5Q8GF0yJFA"

openEI_start <- "https://api.openei.org/utility_rates?version=latest&format=json&sector=Residential&detail=full&approved=true&sort_select=effectiveDate&direction=desc&"
# Set the URL for the job submission endpoint

openEI_url <- paste0(openEI_start, "address=", "46033", "&api_key=", openEI_api_key)
EIresponse <- GET(openEI_url)
# extract contents
content <- content(EIresponse, "text")
utility_data <- fromJSON(content)

 # extract utility rate information
EIutility_df <- list(
  start_date = utility_data$items$startdate,
  #revisions = utility_data$items$revisions,
  label = utility_data$items$label,
  utility = utility_data$items$utility,
  rate_name = utility_data$items$name)

```

```{r}
# Define a vector of zip codes
zipcodes <- msa_zip_pop_est_data
# break up zipcodes to run in openEI  
zipcodes1 <- zipcodes[1:200,]

c(zipcodes1, zipcodes2, zipcodes3, zipcodes4, zipcodes5, zipcodes6)

OpenEI_zip_result <- lapply(zipcodes1$zipcode, get_zip_utility)

test <- get_zip_utility(46033) %>% 
     mutate(start_date = as.Date(start_date/86400, origin="1970-01-01"))

openei_url <- paste0("https://api.openei.org/geocoding?version=latest&format=json&postalcode=", zipcode, "&countrycode=US", "&api_key=U6MEqhK9OJ7obAnmcbzYhpFOenfBM5Q8GF0yJFA")
response <- GET(openei_url)

# Extract the content and parse as JSON
content <- content(response, "text")
json <- fromJSON(content)

```

# Set the payload for the job submission request for reopt
```{r}

zip_utility_info <- get_zip_utility(test$zips)
zip_EI_label <- zip_utility_info$label

test <- one_city_msa_data %>% filter(city == "des moines")

payload <- list(
  "Scenario" = list(
    "Site" = list(
      "address"= test$zips,
      "longitude"= test$lng,
      "latitude" = test$lat,

      "PV" = list(# changes from default
    "module_type" = 1, # premium
    "dc_ac_ratio" = 1.35, # changed default
    "losses" = 0.05,   # changed default
    "federal_itc_pct" = 0.3,  # changed default
     "location" = "roof", # restrict projects to roof
     "can_net_meter" = TRUE
      ),
      "LoadProfile" = list(
        "doe_reference_name" = "MidriseApartment",
        "city" = test$zips
      ),
      "ElectricTariff" = list(
        "urdb_label" = zip_utility_info$label,
        "net_metering_limit_kw" = 1000
      ),
      "Financial" = list(
        "offtaker_tax_pct" = 0, # host effective tax rate
        "owner_tax_pct" = 0.0,
        "co2_cost_us_dollars_per_tonne" = 185
      ))
  )
)

```

```{r STUFF TO WORK ON}
net_metering_limit_kw # by state

```

### Reopt API

```{r}
# Set your API key
api_key <- "GZtZAUcX97VchjrLAeeN3amg2pslaO3hbtCQgXBl"

# Set the URL for the job submission endpoint
url <- "https://developer.nrel.gov/api/reopt/stable/job"

job_url <- paste0(url, api_key)
```

```{r}
# Make the request to submit the job
response <- POST(url, query = list(api_key = api_key), body = payload, encode = "json", timeout(30))

# Check the response status code
status_code <- response$status_code
if (status_code == 200) {
  message("Inputs updated successfully.")
} else {
  message("Error updating inputs.")
}

# get run_uuid to obtain outputs
output <- httr::content(response)
run_uuid <- output$run_uuid
```

```{r}
## GET RESULTS FROM REOPT
results_url <- paste0("https://developer.nrel.gov/api/reopt/stable/job/", run_uuid, "/results")

# Send the GET request to retrieve the results
response <- GET(url = results_url, query = list(api_key = api_key), timeout(30))

# Extract the response content as JSON and convert it to a list
response_content <- content(response, as="text")
response_json <- fromJSON(response_content)

# pull out IMPORTANT data
output_df <- data.frame(
  #IRR = response_json$outputs$Scenario$Site$Financial$irr_pct,
  Health_Emissions_Cost = response_json$outputs$Scenario$Site$lifecycle_emissions_cost_Health_bau,
  Annual_Generation_Renewables = response_json$outputs$Scenario$Site$annual_renewable_electricity_pct)

host/stable/annual_kwh
doe_response <- GET(url = results_url, query = list(api_key = api_key), timeout(30))
```

```{r}
host <- "https://developer.nrel.gov/api/reopt"
api_key <- "?API_KEY=GZtZAUcX97VchjrLAeeN3amg2pslaO3hbtCQgXBl"
version_url <- file.path(host, help_end, api)
api_version <- jsonlite::fromJSON(version_url) # retrieved

```

```{r}
# test scenario post
test <- "https://reopt.nrel.gov/tool/results/f4e6c9db-3c22-43a4-b2bf-15146aec9896"

```